---
title: "DSPS Final Project"
author: "Reese Feldman"
date: "2023-12-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this first part of the analysis, the model is built on "classic" predictors of if someone will vote. 

```{r}
#loading in the data and subsetting by relevant columns and only registered voters
ces2018 <- read.csv("cces18_common_vv(1).csv")
ces2018 <- subset(ces2018, votereg == 1)
ces2018 <- subset(ces2018, select = c(inputstate, birthyr, gender, educ, race, faminc_new, internethome, internetwork, newsint, CC18_308b, CC18_316, CC18_401))
```

```{r}
#reclassify certain variables
ces2018$voted <- ifelse(ces2018$CC18_401 == 5, 1, 0)
ces2018$faminc_new <- ifelse(ces2018$faminc_new == 97, NA, ces2018$faminc_new)
internet <- c(1, 2)
ces2018$hasinternet <- ifelse(ces2018$internethome %in% internet | ces2018$internetwork %in% internet, 1, 0)
ces2018$prevoter <- ifelse(ces2018$CC18_316 == 1, 1, ifelse(ces2018$CC18_316 == 2, 0, NA))
ces2018$age <- 2018 - ces2018$birthyr
#because DC does not have a cost of voting score, we must eliminate rows where the person is from DC
ces2018$inputstate <- ifelse(ces2018$inputstate == 11, NA, ces2018$inputstate)
#rename columns to be more informative
colnames(ces2018)[colnames(ces2018) == "CC18_308b"] = "congapprove"
```

```{r}
#eliminate rows with NA values
ces2018 <- na.omit(ces2018)
```

```{r}
#building a regression model based on typical variables
base2018 <- glm(voted ~ age + gender + race,
                    data = ces2018, family = binomial(link = "logit"))
ces2018$voterlikelihood1 <- predict(base2018, type = "response")
range(ces2018$voterlikelihood1)
```

```{r}
#convert voter likelihood into a prediction
ces2018$predictedvoter1 <- ifelse(ces2018$voterlikelihood1 > .9, 1, 0)
table(actual = ces2018$voted, predicted = ces2018$predictedvoter1)
```

```{r}
## false positive and negative rates
fp1 <- sum(ces2018$predictedvoter1 == 1 & ces2018$voted == 0) / 
  sum(ces2018$voted == 0)
fn1 <- sum(ces2018$predictedvoter1 == 0 & ces2018$voted == 1) / 
  sum(ces2018$voted == 1)
fp1
fn1
```

```{r}
summary(base2018)
```


This next part of the analysis will build the model using additional predictors, in addition to all of the predictors from before

```{r}
#creating the cost of voter index column
covicsv <- read.csv("covi.csv")
covi2018 <- merge(ces2018, covicsv, by = "inputstate")
```

```{r}
#building the regression model with COVI included
new2018 <- glm(voted ~ age + gender + educ + race + faminc_new + hasinternet + COVI + prevoter + newsint + congapprove,
                    data = covi2018, family = binomial(link = "logit"))
covi2018$voterlikelihood2 <- predict(new2018, type = "response")
range(covi2018$voterlikelihood2)
```

```{r}
#convert voter likelihood into a prediction
covi2018$predictedvoter2 <- ifelse(covi2018$voterlikelihood2 > .9, 1, 0)
table(actual = covi2018$voted, predicted = covi2018$predictedvoter2)
```

```{r}
## false positive and negative rates
fp1 <- sum(covi2018$predictedvoter2 == 1 & covi2018$voted == 0) / 
  sum(covi2018$voted == 0)
fn1 <- sum(covi2018$predictedvoter2 == 0 & ces2018$voted == 1) / 
  sum(covi2018$voted == 1)
fp1
fn1
```
```{r}
summary(new2018)
```

```{r}
#building the regression model without race included
best2018 <- glm(voted ~ age + gender + educ + faminc_new + hasinternet + COVI + prevoter + newsint + congapprove,
                    data = covi2018, family = binomial(link = "logit"))
covi2018$voterlikelihood3 <- predict(best2018, type = "response")
range(covi2018$voterlikelihood3)
```

```{r}
#convert voter likelihood into a prediction
covi2018$predictedvoter3 <- ifelse(covi2018$voterlikelihood3 > .9, 1, 0)
table(actual = covi2018$voted, predicted = covi2018$predictedvoter3)
```

```{r}
## false positive and negative rates
fp1 <- sum(covi2018$predictedvoter3 == 1 & covi2018$voted == 0) / 
  sum(covi2018$voted == 0)
fn1 <- sum(covi2018$predictedvoter3 == 0 & ces2018$voted == 1) / 
  sum(covi2018$voted == 1)
fp1
fn1
```

```{r}
summary(best2018)
```


Now that we have built a somewhat good predictive model for voting, we are going to apply the model to the following midterm election in an attempt to predict likely voters "prior" to the 2022 election.

```{r}
ces2022 <- read.csv("CCES22_Common_OUTPUT_vv_topost.csv")
ces2022 <- subset(ces2022, votereg == 1)
ces2022 <- subset(ces2022, select = c(inputstate, birthyr, gender4, educ, faminc_new, internethome, internetwork, newsint, CC22_320b, presvote20post, CC22_401))
```

```{r}
#reclassify certain variables
ces2022$voted <- ifelse(ces2022$CC22_401 == 5, 1, 0)
ces2022$age <- 2022 - ces2022$birthyr
ces2022$faminc_new <- ifelse(ces2022$faminc_new == 97, NA, ces2022$faminc_new)
internet <- c(1, 2)
ces2022$hasinternet <- ifelse(ces2022$internethome %in% internet | ces2022$internetwork %in% internet, 1, 0)
ces2022$prevoter <- ifelse(ces2022$presvote20post == 5, 0, ifelse(is.na(ces2022$presvote20post), NA, 1))
#because DC does not have a cost of voting score, we must eliminate rows where the person is from DC
ces2022$inputstate <- ifelse(ces2022$inputstate == 11, NA, ces2022$inputstate)
#rename columns to be more informative
colnames(ces2022)[colnames(ces2022) == "CC22_320b"] = "congapprove"
colnames(ces2022)[colnames(ces2022) == "gender4"] = "gender"
```

```{r}
#eliminate rows with NA values
ces2022 <- na.omit(ces2022)
```

```{r}
#creating the cost of voter index column
covicsv22 <- read.csv("COVI2022.csv")
colnames(covicsv22)[colnames(covicsv22) == "statenu"] = "inputstate"
covi2022 <- merge(ces2022, covicsv22, by = "inputstate")
```

```{r}
#making predictions for 2022
covi2022$votelikelihood <- predict(best2018, covi2022, type = "response")
range(covi2022$votelikelihood)
```

```{r}
#convert voter likelihood into a prediction
covi2022$predictedvoter22 <- ifelse(covi2022$votelikelihood> .9, 1, 0)
table(actual = covi2022$voted, predicted = covi2022$predictedvoter22)
```

```{r}
## false positive and negative rates
fp1 <- sum(covi2022$predictedvoter22 == 1 & covi2022$voted == 0) / 
  sum(covi2022$voted == 0)
fn1 <- sum(covi2022$predictedvoter22 == 0 & ces2022$voted == 1) / 
  sum(covi2022$voted == 1)
fp1
fn1
```



